---
title: "Coffee Quality Data Cleaning"
output: html_notebook
editor_options: 
  chunk_output_type: inline
---


```{r include=FALSE}
library(tidyverse)
library(xray)
library(dlookr)
library(summarytools)
library(skimr)
library(lubridate)
df_original <- read.csv("df_arabica_clean.csv", header = T)
####https://www.kaggle.com/datasets/fatihb/coffee-quality-data-cqi/download?datasetVersionNumber=1 
```
## Data Problems
```{r message=FALSE, warning=FALSE, echo=FALSE}
print(dfSummary(df_original, 
                style = "grid", plain.ascii = FALSE,
                tmp.img.dir = "./img",
                varnumbers   = FALSE, 
                valid.col    = FALSE, 
                #graph.magnif = 0.76
),
method = 'render')

xray::anomalies(coffee_original)

```

```{r message=FALSE, warning=FALSE}
diagnose_outlier(df_original) %>% 
  filter(outliers_cnt > 0)
```


## Data Cleaning
```{r message=FALSE, warning=FALSE, echo=FALSE}
#remove unnecessary variables
df_clean <- df_original %>% 
    subset(select = -c(X, ID, Farm.Name:Harvest.Year, Status, Defects, Expiration:Certification.Contact)) %>% 
    mutate(Grading.Date = mdy(Grading.Date)) %>%     # convert grading_date from from character to date
    filter(!Processing.Method=='')

# clean up some variable values
df_clean$Variety[is.na(df_clean$Variety) | (df_clean$Variety) == ''] <- "Other"
df_clean$Variety[(df_clean$Variety) =='unknown'] <- "Other"
df_clean$Variety[(df_clean$Variety) =='unknow'] <- "Other"
df_clean$Processing.Method[(df_clean$Processing.Method) =='SEMI-LAVADO'] <- "Semi Washed"
df_clean$Color[(df_clean$Color) =='yello-green' | (df_clean$Color) =='yellow green' | (df_clean$Color) =='yellow- green' ] <- "yellow-green"
df_clean$Color[(df_clean$Color) =='browish-green'] <- "brownish-green"

# check the final result
print(dfSummary(df_clean, 
                style = "grid", plain.ascii = FALSE,
                tmp.img.dir = "./img",
                varnumbers   = FALSE, 
                valid.col    = FALSE, 
                #graph.magnif = 0.76
),
method = 'render')

xray::anomalies(df_clean)
```
1. Dropped variables that are either too messy or irrelevant for the coffee quality analysis (`Farm.Name`, `Mill`, `Harvest.Year`, `ICO.Number`, `Certification.Address`, `Number.of.Bags`, `Bag.Weight`, `Defects`). 
2. Converted `grading_date` to date type.
3. The detected outliers were not removed as the grades can be from 0 to 10, which is within the normal range.
4. Renamed missing values for the `variety` variable and 'unknown'/'unknow' to "Other" variety.
5. Exclude observations with missing `Processing.Method`
6. Group different spellings of "yellow-green" and corrected a typo for brownigh-green color.
7. Although `Category.One.Defects` was highlighted as with an anomaly, it is actually not as its values are boolean and 0 refers to no defects while 1 stands for a defect detected.

```{r echo=FALSE, message=FALSE, warning=FALSE}
## save clean data as new file
df_clean %>% 
    write.csv("coffee_quality_clean.csv", row.names=FALSE)
```

